Integrating brain and biomechanical models - a new paradigm for understanding neuro-muscular control

### 2017-12-05 papers/integration/omm_integ.tex
Figure label refactoring

% Change this - we already talked about channels in the context of the
% three channel model shown. Now talk about how we have a map of
% channels - a continuous space in the \ccg model. This can then lead
% on to Fig 3.
\rvwt{The GPR model in Fig.~\ref{fig:ccg_model}(b) has only three
channels, with the focussed inhibition from striatum to SNr and GPe
defined by a simple one-to-one scheme. The action channels represent
discrete, incompatible motor action choices. In the oculomotor model,
an action channel represents the end-point of saccade, and the
competition carried out in the basal ganglia is between potential
saccade end-points. However, eye movements have a \e{continuous}
end-point space; the eye can rotate to any orientation within its
biomechanically permissible range. Some end-points within this range
are mutually exclusive---it's not possible to look to the left and to
the right simultaneously---but \e{nearby} end-points are not
necessarily incompatible. A small enough error in the end-point of a
saccade will not prevent the eye from foveating on a target as the
foveal region of high visual acuity is not infinitesimally small. To
cope with this requirement, the populations within the oculomotor
basal ganglia are conceived of as two-dimensional topographic grids of
leaky integrator neural elements. Activity in each element corresponds
to a spatial location in the visual field. Neighbouring elements
correspond to locations which are close to each other in the visual
field. Focussed one-to-one projections in the GPR model are replaced
by projective fields with many weighted connections. Specifically,
each unit in Str\_D1 projected to a counterpart SNr$_j$ in SNr with
some weight $w_{max}$, but also connected to neighbouring nodes in SNr
with a weight given by $w_{max}.G(d)$, where $G(d)$ is a circularly
symmetric, 2D-Gaussian which is a function of distance $d$ from
SNr$_j$ [Fig.~\ref{fig:ccg_model}(c)]. A similar scheme applied for
the connectivity from Str\_D2 to GPe and for a number of the other
connections in the \ccg~model; in the SpineML implementation of the
model, this connectivity scheme is named `GaussianKernel'.
% NB - having this Fig. ref here is important for figure ordering!
Fig.~\ref{brain_model} shows a schematic of the SpineML implementation
of the model, based on a diagram output directly from
SpineCreator. Populations for Str\_D1, Str\_D2, STN, SNr and GPe are
shown within the `Basal Ganglia' box. Input comes into the model via
the `World' population and the output population is SC\_deep. Compare
this diagram with Figs.~\ref{fig:ccg_model}(a) \&
(b). Fig.~\ref{brain_model} expands the `SC', `BG' and `slow retinal'
boxes from Fig.~\ref{fig:ccg_model}(a).}

@@

% Retinal model
\rvwt{Input to the \ccg~model is provided through a simple retina
model which directly samples from a larger `world array' of pixel
values. In the current model, the input for the retina is named
`World' and is the retinotopic projection [Fig.~\ref{brain_model}(b)]
of the eye's field of view of the world [Fig.~\ref{brain_model}(a)]
and the luminant targets therein. The raw input in `World' is fed into
a population which adds noise, and then via a delayed connection to
FEF (the sustained retinal input path), to simulate processing through
the dorsal visual stream. It is also fed, without substantial delay,
into two leaky integrator layers (Retina\_1 and Retina\_2) with
different time constants, with the more slowly reacting layer
(Retina\_2) inhibiting its faster counterpart. The faster layer
responds quickly to the appearance of a prolonged stimulus before it
is inhibited by the slow layer, forming a phasic response to stimulus
onset. The mechanism ensures that phasic rather than tonic responses
arrive at the superficial SC from the retina.}

@@

Briefly, the model consists of input from the World population (see
Fig.~\ref{brain_model}, green population box) producing activity in an
`express' pathway to superior colliculus (purple) and simultaneously
in cortex, represented here by the FEF population (grey boxes in
Fig.~\ref{brain_model}). The express pathway causes short latency
activity in the superficial superior colliculus, which directly
innervates the deeper layers of the superior colliculus
(SC\_deep). Activity in FEF generates firing in a
thalamo-cortico-basal ganglia loop. The output of the basal ganglia is
the substantia nigra pars reticulata (SNr) which tonically inhibits
SC\_deep. If a location of activity in FEF is able to dominate
selection in the basal ganglia circuit, the corresponding location in
SNr will dis-inhibit and activity will build up in SC\_deep encoding
the saccade end point.

@@

In addition to the brain model components, all of which are
code-generated using \stob~as described above, two hand-written
components are integrated into the model: The biomechanical eye model
and a sensory input component. The sensory input component takes the
eye's rotational state and the state of the experimental luminances
and projects a retinotopic activity map into the brain model. Both of
these BRAHMS components were hand-written in C++. To incorporate these
components into the SpineML model, a \stob~\e{external.xsl} file was
used.  The external.xsl file scheme for incorporating external BRAHMS
components into a SpineML model was a new \stob~feature motivated by
the current work.  Fig.~\ref{model_framework} shows the workflow, in
which the model specification files (blue box - a combination of
SpineML files and C++ code), are processed (green box) into a BRAHMS
system (red box).

@@

This led us to hypothesise that the retinotopic mapping \rvwt{to the
SBG be preceded} by an associated widening projection field such that
the hill of activity in \rvwt{a `final' deep layer of superior
colliculus is invariant with position on the map}. There are a number of
locations in the system in which this widening projection field could
exist. It could be implemented in the projections between the retinal
populations and the superficial layer of SC along with the projection
between the World and the FEF population. However, this would affect
activity within the basal ganglia of the model, contradicting a result
in \cite{cope_basal_2017} which explains the `hockey stick' profile
for saccade latency as a function of saccade eccentricity. Instead, we
suggest that a widening projection field is encoded within the
superior colliculus itself, a complex, multi-layered structure which
could quite plausibly support such a function. Indeed, such widening
activity can be seen in the stimulation experiments
in \cite{vokoun_intralaminar_2010} and \cite{vokoun_response_2014}.
%
\rvw{In \cite{ghitani_excitatory_2014}, from the same research group,
evidence is presented for an excitatory and widely projecting pathway
from the stratum griseum intermediale (equivalent to our SC\_deep) to
the more superficial layers stratum opticum and stratum griseum
superficiale. Although this pathway is a `wide' projecting field, the
experiments do not indicate whether the projection \e{widens} along
the rostral-caudal axis of the SC. \cite{bayguinov_hard-wired_2015}
presents evidence for another projecting field within SC whose
connectivity pattern \e{does} change along the rostral-caudal
axis. This projection is inhibitory in nature. Although neither of
these results precisely match the widening, excitatory projection
field hypothesised here, they do indicate that such connection
patterns are plausible.}
%
Although in this work we do not model
the SC in detail, we extended the model with a third functional layer
named SC\_deep2, shown in Fig.~\ref{scdeep}(b) (\ccg~has only the two
layers SC\_sup and SC\_deep). We introduced a widening projection
based on a Gaussian projection field whose width, $\sigma(r)$ varies
in inverse proportion to the magnification factor, $M(r)$, given in
Eq.~\ref{eq:cmf} according to:
\begin{equation} \label{eq:sigmar}
\sigma(r) = \frac{m_{\sigma}}{M(r)} - \frac{m_{\sigma}}{M^0} + \sigma_0 \qquad r > r_0
\end{equation}
$m_{\sigma}$ is a scalar parameter which determines the `magnitude of
the widening'. $M^0$ is the `starting' magnification factor; within
the foveal region ($0 \leq r \leq r_0$), the projection field is not
allowed to widen and so
\begin{equation} \label{eq:sigmar2}
\sigma(r) = \sigma_0 \qquad r \leq r_0
\end{equation}
which makes $\sigma_0$ the width of the Gaussian projection field
within the foveal region. (Note that the value chosen for the width of
the foveal region, $r_0$ is not identical to the foveal shift
parameter used in the \e{DecayingAtFovea} projections into striatum.)
The \e{Widening Gaussian} projection weight, $w(r,d)$ is then computed
as:
\begin{equation} \label{eq:widening}
w(r,d) = e^{-\frac{d^2}{2\sigma\left(r\right)^2}}
\end{equation}
where $d$ is the distance between the source and destination elements
in the collicular plane.  $m_\sigma$ was set to 50, $\sigma_0$ was
0.3, $M^0$ was 12.43 and $r_0$ was 20.
% Could have a figure here, made from widening_gaussian.m

@@

We found the best parameters for the exponential in
Eq.~\ref{eq:weightmaps} ($i$ and $j$) by a manual tuning
process. After selecting values for $i$ and $j$ in either the
horizontal or vertical/oblique channels, we ran the model 6 times at
each of 8 target eccentricities (7\dg--14\dg) which were purely in the
direction of the newly parameterised channel. The training saccades
were produced as described below in Sect.~\ref{sec:results:singlesaccades},
with the same fixation and target luminances (crosses of magnitude 0.2
and 0.3) but with the fixation offset and target onset occurring at
0.2~s. We measured the end-point of the saccade by detecting the
location at which the saccade velocity had dropped below 0.005 of its
peak. We iterated until the mean saccade endpoint plotted versus
target was close to the ideal straight line---see
Fig.~\ref{sacc_vs_targ}(a) \& \rvw{(b)}. We applied the same parameters to
both directions of each channel; $i_{up} = i_{down} = 0.00195$,
$j_{up} = j_{down} = 0.075$, $i_{left} = i_{right} = 0.0016$ and
$j_{left} = j_{right} = 0.067$.

The resulting weight maps (where the oblique maps are
1/10$^{\mathrm{th}}$ of the vertical maps, as described earlier) are
shown in Fig.~\ref{weightmaps}.
%
\rvw{First, recall that the $r$ axis of the neural surface corresponds
to the amplitude of a saccade and the $\phi$ axis indicates the polar
direction of the saccade, as described in Sect.~\ref{sec:mapping} and
Fig.~\ref{fig:mapping}. Fig.~\ref{weightmaps}(a) shows the weight map
for the muscle which rotates the eye to the left. As we modelled a
left eye, this actuates the lateral rectus muscle. The exponential
rise of Eq.~\ref{eq:weightmaps} is seen in the $r$ direction; as $r$
increases, so the connection strength to the SBG channel rises
exponentially. The connection strength is greatest along the centre
line, for a value of $\phi$ which corresponds to a purely leftward
movement. Note that $\phi$ is presented in neural coordinates, and not
in degrees or radians; 1$\leq\phi\leq$50 corresponds to a range of
0\dg~to 360\dg; $\phi=$38.5 corresponds to movements left. The
connections strength drops away sinusoidally as $\phi$ moves away from
the centre line at $\phi=$38.5. In regions of the map for which there
is no leftward movement, that is, in the half of the map which
corresponds to any movement with a rightward component, the `left'
weight map is 0. Fig.~\ref{weightmaps}(d) shows the weight map for
rightward movements, actuating the medial rectus muscle of the
eye. The line of maximum connection strength is along $\phi=$13.5. The
map is a mirror of Fig.~\ref{weightmaps}(a), reflected about the line
$\phi$=26. The `left' and `right' weight maps are orthogonal; the
non-zero region of the `left' map is zero in the `right' map and vice
versa. Fig~\ref{weightmaps}(b) \& (d) show the weight maps for
downward and upward eye movements; the `down' map activates the SBG
channel for the inferior rectus muscle, the `up' map activates the
superior rectus. Note that `down' is not orthogonal either to `left'
or `right' because a saccade down and left is achieved by
simultaneously activating both the lateral and inferior rectus
muscles. However, the `up' map \e{is} orthogonal to the `down' map and
spans the edges of the surface where $\phi$ rolls over from 1 to
50. The line of maximum connection strength for the `up' map is along
$\phi=$1; for `down' $\phi=$26. Based on the training described in
Sect.~\ref{sec:methods:integrating}, the maps driving the superior
oblique (`Z+') and inferior oblique (`Z-') muscles were set to
1/10$^{\mathrm{th}}$ of the `down' and `up' maps.}
% From map_modelling2.m:
%    yh=0.0016*exp(0.067.*x);
%    yv=0.00195*exp(0.075.*x);

@@

In Fig.~\ref{sacc_vs_targ}, we showed the result of running the model
to targets located on the principle axes, on which the model was
trained. We then simulated single saccades to targets in one hemifield
of the eye's field of view, with eccentricities between 6\dg~and
14.5\dg. As in the training, we ran the simulation 6 times for each
target, $\mathbf{\theta}^t = (\theta_{x}^t, \theta_{y}^t, 0)$ to
obtain mean saccade end-points.  Fig.~\ref{errorsurfaceTM3} shows
saccade accuracy results for an entire hemifield in the na\"ive model
which passed the output of SC\_deep directly to SBG via the weight
maps. The ratio of the magnitude of the error vector to the magnitude
of the target vector is plotted using a colour map. This ratio is
shown for the full, three dimensional error vector in
Fig.~\ref{errorsurfaceTM3}(a) and for the $x$, $y$ and $z$ components
in Figs.~\ref{errorsurfaceTM3}(b)--(c).  Inspection of
Fig.~\ref{errorsurfaceTM3}(a) shows that the end-point error is
minimal along the principle axes ($\theta_{x}^t=\;$0 or
$\theta_{y}^t=\;$0) and maximal near the 45\dg~oblique targets (blue
lines) with the end point error as high as 80\% of the programmed
saccade magnitude.  The $x$ component error map in
Fig.~\ref{errorsurfaceTM3}(b) shows the same trend, mirrored about the
`Target X' axis, whereas the $y$ and $z$ component errors are,
relatively, much smaller.  Because the $x$ component of the error is
clearly contributing to end point errors which would not be considered
`on target', especially for oblique saccades, we considered the effect
of the non-uniform size of the hill of activity in SC\_deep.

@@

To bring our model in line with these results, whilst maintaining the
eccentricity dependent activity in basal ganglia, we hypothesised that
a `widening projection' exists between two maps in superior
colliculus. \rvw{As described in
Sect.~\ref{sec:methods:integrating}, there is
now experimental evidence for similar projections
\citep{ghitani_excitatory_2014,bayguinov_hard-wired_2015} making this
a plausible suggestion.}  Activities in one SC\_deep layer remains
eccentricity-dependent, with loops back to thalamus and cortex and
through basal ganglia.  This activity is then fed through a
projection, which applies a Gaussian projection field, whose width
increases with increasing stimulus eccentricity according to
Eq.~\ref{eq:widening}.  The activity in this second SC\_deep layer is
then fed to the weight maps of the SBG. This model was called
`TModel4'. TModel4 was parameterised such that its horizontal and
vertical error was similar---so that its equivalent of
Fig.~\ref{sacc_vs_targ} showed a similar sum of squares error.

Figs.~\ref{errorsurface}(a)--(d) show the \rvw{percentage errors for
TModel4}.  First of
all, note that the error magnitudes are much smaller. The mean errors
are smaller for every axis. The largest errors produced by the model
are approximately 15\%, which are within the boundaries of what some
authors have suggested would be regarded as an accurate
saccade \citep{mcpeek_saccade_2002,mcpeek_incomplete_2006}.  The
magnitude of the largest error vector is approximately 1.5\dg.

@@

Having finalised the model by setting the weight maps, we then
proceeded to exercise the model (TModel4), starting with saccades to a
single target; prosaccades. Fig.~\ref{outmany}(a) shows 9
representative saccades to a single target luminance. Initially, the
eye had rotational state $\theta_x=\theta_y=\theta_z=0$ with its fovea
directed at a fixation luminance cross (span 6\dg, bar width 2\dg) of
magnitude 0.2 (in arbitrary units). At a simulation time of 0.4~s, the
fixation luminance was set to 0 and a target luminance cross of the
same dimensions as the fixation but with magnitude 0.3 was illuminated
at one of the 9 different locations, marked by crosses in
Fig.~\ref{outmany}(a).  The resulting trajectories are plotted, with
colour indicating the relationship between trajectories and target
crosses. The approximate end-point error is visible in this figure,
although the last point in each trajectory is the saccade position at
0.8~s and not the velocity-based end-point described
above. Figs.~\ref{outmany}(b) and (c) show the rotational components
of the blue and red trajectories in Fig.~\ref{outmany}(a) along with
the target and fixation luminance values. Rotations are the eye's
Euler rotational components in the world frame of reference.

@@

To verify that our implementation of the brain model has the same
functionality as that reported in \cite{cope_basal_2017}, we
investigated the effect on saccadic response times of: target
eccentricity; and any gap or overlap between fixation off-time and
target on-time. We showed that the full model reproduces the `hockey
stick' shape shown in Fig. 7 of \cite{cope_basal_2017} for horizontal
[Fig.~\ref{lat_vs_all}(a)], vertical [Fig.~\ref{lat_vs_all}(b)] and
oblique saccades (not shown). The latency increases with eccentricity
far from the fovea because the retinotopic mapping reduces the
activity in the basal ganglia for more eccentric targets (see
Fig.~\ref{fig:mapping}). Closer to the fovea, the \rvwt{effect of} the
foveal mask \rvwt{on} the activity in FEF again leads to reduced input
into the basal ganglia and an increased time to achieve disinhibition
in SNr.

>>>>

% Change this - we already talked about channels in the context of the
% three channel model shown. Now talk about how we have a map of
% channels - a continuous space in the \ccg model. This can then lead
% on to Fig 3.
\rvwt{The GPR model in Fig.~\ref{fig:ccg_model}(b) has only three
channels, with the focussed inhibition from striatum to SNr and GPe
defined by a simple one-to-one scheme. The action channels represent
discrete, incompatible motor action choices. In the oculomotor model,
an action channel represents the end-point of saccade, and the
competition carried out in the basal ganglia is between potential
saccade end-points. However, eye movements have a \e{continuous}
end-point space; the eye can rotate to any orientation within its
biomechanically permissible range. Some end-points within this range
are mutually exclusive---it's not possible to look to the left and to
the right simultaneously---but \e{nearby} end-points are not
necessarily incompatible. A small enough error in the end-point of a
saccade will not prevent the eye from foveating on a target as the
foveal region of high visual acuity is not infinitesimally small. To
cope with this requirement, the populations within the oculomotor
basal ganglia are conceived of as two-dimensional topographic grids of
leaky integrator neural elements. Activity in each element corresponds
to a spatial location in the visual field. Neighbouring elements
correspond to locations which are close to each other in the visual
field. Focussed one-to-one projections in the GPR model are replaced
by projective fields with many weighted connections. Specifically,
each unit in Str\_D1 projected to a counterpart SNr$_j$ in SNr with
some weight $w_{max}$, but also connected to neighbouring nodes in SNr
with a weight given by $w_{max}.G(d)$, where $G(d)$ is a circularly
symmetric, 2D-Gaussian which is a function of distance $d$ from
SNr$_j$ [Fig.~\ref{fig:ccg_model}(c)]. A similar scheme applied for
the connectivity from Str\_D2 to GPe and for a number of the other
connections in the \ccg~model; in the SpineML implementation of the
model, this connectivity scheme is named `GaussianKernel'.
% NB - having this Fig. ref here is important for figure ordering!
Fig.~\ref{fig:brainmodel} shows a schematic of the SpineML implementation
of the model, based on a diagram output directly from
SpineCreator. Populations for Str\_D1, Str\_D2, STN, SNr and GPe are
shown within the `Basal Ganglia' box. Input comes into the model via
the `World' population and the output population is SC\_deep. Compare
this diagram with Figs.~\ref{fig:ccg_model}(a) \&
(b). Fig.~\ref{fig:brainmodel} expands the `SC', `BG' and `slow retinal'
boxes from Fig.~\ref{fig:ccg_model}(a).}

@@

% Retinal model
\rvwt{Input to the \ccg~model is provided through a simple retina
model which directly samples from a larger `world array' of pixel
values. In the current model, the input for the retina is named
`World' and is the retinotopic projection [Fig.~\ref{fig:brainmodel}(b)]
of the eye's field of view of the world [Fig.~\ref{fig:brainmodel}(a)]
and the luminant targets therein. The raw input in `World' is fed into
a population which adds noise, and then via a delayed connection to
FEF (the sustained retinal input path), to simulate processing through
the dorsal visual stream. It is also fed, without substantial delay,
into two leaky integrator layers (Retina\_1 and Retina\_2) with
different time constants, with the more slowly reacting layer
(Retina\_2) inhibiting its faster counterpart. The faster layer
responds quickly to the appearance of a prolonged stimulus before it
is inhibited by the slow layer, forming a phasic response to stimulus
onset. The mechanism ensures that phasic rather than tonic responses
arrive at the superficial SC from the retina.}

@@

Briefly, the model consists of input from the World population (see
Fig.~\ref{fig:brainmodel}, green population box) producing activity in an
`express' pathway to superior colliculus (purple) and simultaneously
in cortex, represented here by the FEF population (grey boxes in
Fig.~\ref{fig:brainmodel}). The express pathway causes short latency
activity in the superficial superior colliculus, which directly
innervates the deeper layers of the superior colliculus
(SC\_deep). Activity in FEF generates firing in a
thalamo-cortico-basal ganglia loop. The output of the basal ganglia is
the substantia nigra pars reticulata (SNr) which tonically inhibits
SC\_deep. If a location of activity in FEF is able to dominate
selection in the basal ganglia circuit, the corresponding location in
SNr will dis-inhibit and activity will build up in SC\_deep encoding
the saccade end point.

@@

In addition to the brain model components, all of which are
code-generated using \stob~as described above, two hand-written
components are integrated into the model: The biomechanical eye model
and a sensory input component. The sensory input component takes the
eye's rotational state and the state of the experimental luminances
and projects a retinotopic activity map into the brain model. Both of
these BRAHMS components were hand-written in C++. To incorporate these
components into the SpineML model, a \stob~\e{external.xsl} file was
used.  The external.xsl file scheme for incorporating external BRAHMS
components into a SpineML model was a new \stob~feature motivated by
the current work.  Fig.~\ref{fig:modelframework} shows the workflow, in
which the model specification files (blue box - a combination of
SpineML files and C++ code), are processed (green box) into a BRAHMS
system (red box).

@@

This led us to hypothesise that the retinotopic mapping \rvwt{to the
SBG be preceded} by an associated widening projection field such that
the hill of activity in \rvwt{a `final' deep layer of superior
colliculus is invariant with position on the map}. There are a number of
locations in the system in which this widening projection field could
exist. It could be implemented in the projections between the retinal
populations and the superficial layer of SC along with the projection
between the World and the FEF population. However, this would affect
activity within the basal ganglia of the model, contradicting a result
in \cite{cope_basal_2017} which explains the `hockey stick' profile
for saccade latency as a function of saccade eccentricity. Instead, we
suggest that a widening projection field is encoded within the
superior colliculus itself, a complex, multi-layered structure which
could quite plausibly support such a function. Indeed, such widening
activity can be seen in the stimulation experiments
in \cite{vokoun_intralaminar_2010} and \cite{vokoun_response_2014}.
%
\rvw{In \cite{ghitani_excitatory_2014}, from the same research group,
evidence is presented for an excitatory and widely projecting pathway
from the stratum griseum intermediale (equivalent to our SC\_deep) to
the more superficial layers stratum opticum and stratum griseum
superficiale. Although this pathway is a `wide' projecting field, the
experiments do not indicate whether the projection \e{widens} along
the rostral-caudal axis of the SC. \cite{bayguinov_hard-wired_2015}
presents evidence for another projecting field within SC whose
connectivity pattern \e{does} change along the rostral-caudal
axis. This projection is inhibitory in nature. Although neither of
these results precisely match the widening, excitatory projection
field hypothesised here, they do indicate that such connection
patterns are plausible.}
%
Although in this work we do not model
the SC in detail, we extended the model with a third functional layer
named SC\_deep2, shown in Fig.~\ref{fig:scdeep}(b) (\ccg~has only the two
layers SC\_sup and SC\_deep). We introduced a widening projection
based on a Gaussian projection field whose width, $\sigma(r)$ varies
in inverse proportion to the magnification factor, $M(r)$, given in
Eq.~\ref{eq:cmf} according to:
\begin{equation} \label{eq:sigmar}
\sigma(r) = \frac{m_{\sigma}}{M(r)} - \frac{m_{\sigma}}{M^0} + \sigma_0 \qquad r > r_0
\end{equation}
$m_{\sigma}$ is a scalar parameter which determines the `magnitude of
the widening'. $M^0$ is the `starting' magnification factor; within
the foveal region ($0 \leq r \leq r_0$), the projection field is not
allowed to widen and so
\begin{equation} \label{eq:sigmar2}
\sigma(r) = \sigma_0 \qquad r \leq r_0
\end{equation}
which makes $\sigma_0$ the width of the Gaussian projection field
within the foveal region. (Note that the value chosen for the width of
the foveal region, $r_0$ is not identical to the foveal shift
parameter used in the \e{DecayingAtFovea} projections into striatum.)
The \e{Widening Gaussian} projection weight, $w(r,d)$ is then computed
as:
\begin{equation} \label{eq:widening}
w(r,d) = e^{-\frac{d^2}{2\sigma\left(r\right)^2}}
\end{equation}
where $d$ is the distance between the source and destination elements
in the collicular plane.  $m_\sigma$ was set to 50, $\sigma_0$ was
0.3, $M^0$ was 12.43 and $r_0$ was 20.
% Could have a figure here, made from widening_gaussian.m

@@

We found the best parameters for the exponential in
Eq.~\ref{eq:weightmaps} ($i$ and $j$) by a manual tuning
process. After selecting values for $i$ and $j$ in either the
horizontal or vertical/oblique channels, we ran the model 6 times at
each of 8 target eccentricities (7\dg--14\dg) which were purely in the
direction of the newly parameterised channel. The training saccades
were produced as described below in Sect.~\ref{sec:results:singlesaccades},
with the same fixation and target luminances (crosses of magnitude 0.2
and 0.3) but with the fixation offset and target onset occurring at
0.2~s. We measured the end-point of the saccade by detecting the
location at which the saccade velocity had dropped below 0.005 of its
peak. We iterated until the mean saccade endpoint plotted versus
target was close to the ideal straight line---see
Fig.~\ref{fig:saccvstarg}(a) \& \rvw{(b)}. We applied the same parameters to
both directions of each channel; $i_{up} = i_{down} = 0.00195$,
$j_{up} = j_{down} = 0.075$, $i_{left} = i_{right} = 0.0016$ and
$j_{left} = j_{right} = 0.067$.

The resulting weight maps (where the oblique maps are
1/10$^{\mathrm{th}}$ of the vertical maps, as described earlier) are
shown in Fig.~\ref{fig:weightmaps}.
%
\rvw{First, recall that the $r$ axis of the neural surface corresponds
to the amplitude of a saccade and the $\phi$ axis indicates the polar
direction of the saccade, as described in Sect.~\ref{sec:mapping} and
Fig.~\ref{fig:mapping}. Fig.~\ref{fig:weightmaps}(a) shows the weight map
for the muscle which rotates the eye to the left. As we modelled a
left eye, this actuates the lateral rectus muscle. The exponential
rise of Eq.~\ref{eq:weightmaps} is seen in the $r$ direction; as $r$
increases, so the connection strength to the SBG channel rises
exponentially. The connection strength is greatest along the centre
line, for a value of $\phi$ which corresponds to a purely leftward
movement. Note that $\phi$ is presented in neural coordinates, and not
in degrees or radians; 1$\leq\phi\leq$50 corresponds to a range of
0\dg~to 360\dg; $\phi=$38.5 corresponds to movements left. The
connections strength drops away sinusoidally as $\phi$ moves away from
the centre line at $\phi=$38.5. In regions of the map for which there
is no leftward movement, that is, in the half of the map which
corresponds to any movement with a rightward component, the `left'
weight map is 0. Fig.~\ref{fig:weightmaps}(d) shows the weight map for
rightward movements, actuating the medial rectus muscle of the
eye. The line of maximum connection strength is along $\phi=$13.5. The
map is a mirror of Fig.~\ref{fig:weightmaps}(a), reflected about the line
$\phi$=26. The `left' and `right' weight maps are orthogonal; the
non-zero region of the `left' map is zero in the `right' map and vice
versa. Fig~\ref{fig:weightmaps}(b) \& (d) show the weight maps for
downward and upward eye movements; the `down' map activates the SBG
channel for the inferior rectus muscle, the `up' map activates the
superior rectus. Note that `down' is not orthogonal either to `left'
or `right' because a saccade down and left is achieved by
simultaneously activating both the lateral and inferior rectus
muscles. However, the `up' map \e{is} orthogonal to the `down' map and
spans the edges of the surface where $\phi$ rolls over from 1 to
50. The line of maximum connection strength for the `up' map is along
$\phi=$1; for `down' $\phi=$26. Based on the training described in
Sect.~\ref{sec:methods:integrating}, the maps driving the superior
oblique (`Z+') and inferior oblique (`Z-') muscles were set to
1/10$^{\mathrm{th}}$ of the `down' and `up' maps.}
% From map_modelling2.m:
%    yh=0.0016*exp(0.067.*x);
%    yv=0.00195*exp(0.075.*x);

@@

In Fig.~\ref{fig:saccvstarg}, we showed the result of running the model
to targets located on the principle axes, on which the model was
trained. We then simulated single saccades to targets in one hemifield
of the eye's field of view, with eccentricities between 6\dg~and
14.5\dg. As in the training, we ran the simulation 6 times for each
target, $\mathbf{\theta}^t = (\theta_{x}^t, \theta_{y}^t, 0)$ to
obtain mean saccade end-points.  Fig.~\ref{fig:errorsurfaceTM3} shows
saccade accuracy results for an entire hemifield in the na\"ive model
which passed the output of SC\_deep directly to SBG via the weight
maps. The ratio of the magnitude of the error vector to the magnitude
of the target vector is plotted using a colour map. This ratio is
shown for the full, three dimensional error vector in
Fig.~\ref{fig:errorsurfaceTM3}(a) and for the $x$, $y$ and $z$ components
in Figs.~\ref{fig:errorsurfaceTM3}(b)--(c).  Inspection of
Fig.~\ref{fig:errorsurfaceTM3}(a) shows that the end-point error is
minimal along the principle axes ($\theta_{x}^t=\;$0 or
$\theta_{y}^t=\;$0) and maximal near the 45\dg~oblique targets (blue
lines) with the end point error as high as 80\% of the programmed
saccade magnitude.  The $x$ component error map in
Fig.~\ref{fig:errorsurfaceTM3}(b) shows the same trend, mirrored about the
`Target X' axis, whereas the $y$ and $z$ component errors are,
relatively, much smaller.  Because the $x$ component of the error is
clearly contributing to end point errors which would not be considered
`on target', especially for oblique saccades, we considered the effect
of the non-uniform size of the hill of activity in SC\_deep.

@@

To bring our model in line with these results, whilst maintaining the
eccentricity dependent activity in basal ganglia, we hypothesised that
a `widening projection' exists between two maps in superior
colliculus. \rvw{As described in
Sect.~\ref{sec:methods:integrating}, there is
now experimental evidence for similar projections
\citep{ghitani_excitatory_2014,bayguinov_hard-wired_2015} making this
a plausible suggestion.}  Activities in one SC\_deep layer remains
eccentricity-dependent, with loops back to thalamus and cortex and
through basal ganglia.  This activity is then fed through a
projection, which applies a Gaussian projection field, whose width
increases with increasing stimulus eccentricity according to
Eq.~\ref{eq:widening}.  The activity in this second SC\_deep layer is
then fed to the weight maps of the SBG. This model was called
`TModel4'. TModel4 was parameterised such that its horizontal and
vertical error was similar---so that its equivalent of
Fig.~\ref{fig:saccvstarg} showed a similar sum of squares error.

Figs.~\ref{fig:errorsurface}(a)--(d) show the \rvw{percentage errors for
TModel4}.  First of
all, note that the error magnitudes are much smaller. The mean errors
are smaller for every axis. The largest errors produced by the model
are approximately 15\%, which are within the boundaries of what some
authors have suggested would be regarded as an accurate
saccade \citep{mcpeek_saccade_2002,mcpeek_incomplete_2006}.  The
magnitude of the largest error vector is approximately 1.5\dg.

@@

Having finalised the model by setting the weight maps, we then
proceeded to exercise the model (TModel4), starting with saccades to a
single target; prosaccades. Fig.~\ref{fig:outmany}(a) shows 9
representative saccades to a single target luminance. Initially, the
eye had rotational state $\theta_x=\theta_y=\theta_z=0$ with its fovea
directed at a fixation luminance cross (span 6\dg, bar width 2\dg) of
magnitude 0.2 (in arbitrary units). At a simulation time of 0.4~s, the
fixation luminance was set to 0 and a target luminance cross of the
same dimensions as the fixation but with magnitude 0.3 was illuminated
at one of the 9 different locations, marked by crosses in
Fig.~\ref{fig:outmany}(a).  The resulting trajectories are plotted, with
colour indicating the relationship between trajectories and target
crosses. The approximate end-point error is visible in this figure,
although the last point in each trajectory is the saccade position at
0.8~s and not the velocity-based end-point described
above. Figs.~\ref{fig:outmany}(b) and (c) show the rotational components
of the blue and red trajectories in Fig.~\ref{fig:outmany}(a) along with
the target and fixation luminance values. Rotations are the eye's
Euler rotational components in the world frame of reference.

@@

To verify that our implementation of the brain model has the same
functionality as that reported in \cite{cope_basal_2017}, we
investigated the effect on saccadic response times of: target
eccentricity; and any gap or overlap between fixation off-time and
target on-time. We showed that the full model reproduces the `hockey
stick' shape shown in Fig. 7 of \cite{cope_basal_2017} for horizontal
[Fig.~\ref{fig:latvsall}(a)], vertical [Fig.~\ref{fig:latvsall}(b)] and
oblique saccades (not shown). The latency increases with eccentricity
far from the fovea because the retinotopic mapping reduces the
activity in the basal ganglia for more eccentric targets (see
Fig.~\ref{fig:mapping}). Closer to the fovea, the \rvwt{effect of} the
foveal mask \rvwt{on} the activity in FEF again leads to reduced input
into the basal ganglia and an increased time to achieve disinhibition
in SNr.

Fig.~\ref{fig:latvsall}(c) shows latencies achieved when varying the
time between fixation offset and target onset. This is termed
the \e{gap condition}; and is represented by a scalar value which, if
positive, refers to a gap between fixation offset and target onset,
and when negative, signifies an overlap, with the fixation luminance
persisting past the time at which the target is illuminated. A
negative gap is also termed an \e{overlap}.  Again, we verify the
behaviour presented in \cite{cope_basal_2017}, explained as resulting
from the inhibition of the cortico-thalamic loop by SNr. In the gap
condition, when the fixation luminance is removed, activity in STN
immediately begins to decay, allowing SNr activity to reduce and
thereby reducing inhibition on thalamus, allowing the target luminance
to build up quickly in FEF, thalamus and through the basal ganglia's
striatum and SNr. The shape of the curves in Fig.~\ref{fig:latvsall}(c)
matches the results in \cite{cope_basal_2017} for target luminances of
1 and 0.6; for overlaps longer than 100~ms (gap $<$ -100~ms), the
latency becomes constant; the saccade is programmed whilst the
fixation is present, with the target luminance inducing sufficient
activity in striatum to `break through' the SNr inhibition caused by
the fixation. If the target luminance is reduced to 0.3, the balance
is altered in favour of the fixation and the latency vs.~gap becomes
approximately linear and equal to the overlap time plus around 100~ms.

Fig.~\ref{fig:latvsall}(d) shows the effect of the dopamine parameter on
saccade latencies in gap, step and overlap conditions. In general, the
effect of decreasing the dopamine parameter was a smooth, monotonic
and undramatic increase in saccade latency. However, the data for the
overlap condition with a target luminance which was 3 times as bright
as the fixation luminance was more interesting. Here we see a
transition around a dopamine value if 0.7.  Below this value, the
basal ganglia is not able to select the target luminance until the
fixation is removed, reducing the excitatory drive from STN to SNr,
and consequently the inhibition from SNr to the thalamo-cortical
loop. For the target luminance 0.6, 0.7 dopamine allows the basal
ganglia to select sufficiently well so that the target can build up in
the thalamo-cortical loop, in spite of the fixation
overlap.
